---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidygedcom

<!-- badges: start -->
<!-- badges: end -->

Import, create, summarise, and edit family tree GEDCOM files using tidy dataframes. This package is still in heavy development with the first operational version unlikely to be complete before October 2020.

## Installation

You can install development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jl5000/tidygedcom")
```

## Example

The intent is to allow the user to import GEDCOM files or create them from scratch using a `ggplot2` type interface; starting with a base GEDCOM object (with only a header and trailer section) and adding records as you would add layers to a ggplot object (but instead using the pipe operator, %>%).

An example is below:

```{r example}
library(tidygedcom)

tg <- gedcom(subm("Jamie Lendrum")) %>%
  add_individual(sex = "M", individual_notes = "The central character in the Star Wars Skywalker Saga") %>%
  add_individual_names("Anakin Skywalker", type = "birth", given = "Anakin", surname = "Skywalker") %>% 
  add_individual_event(event_type = "DEAT", age_at_event = "45y", cause_of_event = "Killed by son Luke",
                       place_name = "Second Death Star", place_notes = "Orbiting Endor System") %>% 
  add_individual_attribute(attribute_type = "RELI", attribute_descriptor = "Jedi") %>% 
  add_individual_attribute(attribute_type = "PROP", attribute_descriptor = "Lightsaber") %>%
  add_individual(sex = "F", individual_notes = "Queen of Naboo") %>%
  add_individual_names("Padme Amidala") %>% 
  add_individual(sex = "F") %>% 
  add_individual_names("Leia Organa") %>% 
  add_individual(sex = "M") %>%
  add_individual_names("Luke Skywalker") %>% 
  add_family(husband = "Anakin", wife = "Padme", children = c("Luke", "Leia"), submitters = "Jamie") %>%
  add_individual(sex = "M") %>% 
  add_individual_names("Obi-Wan Kenobi") %>% 
  activate_individual_record("Anakin") %>% 
  add_individual_association("Obi-Wan", association = "Master") %>% 
  add_note("Based on Star Wars") %>% 
  add_source(short_title = "Star Wars", title = "Star Wars Episode IV: A New Hope") %>% 
  add_repository("The Skywalker Saga") %>% 
  add_submitter("George Lucas") %>%
  add_multimedia(file_reference = "XYZ", format = "jpg")

tg
```


Just like a ggplot object requires aesthetics, a GEDCOM file requires details of a submitter. If no submitter details are given, the username is used.

Within the package, GEDCOM files are represented as tidy tibbles (a sub-class of tibble known as `tidygedcom` objects) allowing easy manipulation and exploitation of existing `tidyverse` infrastructure.

Functionality will also exist to allow the removal and (limited) modification of existing records, as well as summaries:

```{r}
num_indi(tg)
num_fam(tg)

str(tg)
summary(tg)

df_individuals(tg)
df_families(tg)
```

Editing of existing records is made possible through 'activation' (much like the `tidygraph` package). When a record is created, it automatically becomes the active record, through an object attribute. Record editing functions then act on this record. Other records can be activated through a series of `activate_*_record()` functions.

A sister package named `tgvis` will allow visualisation of `tidygedcom` objects.

## Current status

* The GEDCOM 5.5.1 specification has been implemented within the params.R, structures.R, records.R, and validate_params.R files. This is the bulk of the internal data structures required;
* Some initial functions to summarise tidygedcom objects have been completed;
* Functions to import, export and create tidygedcom objects have been completed;
* Current work is on the modification of tidygedcom objects;
* It is still undecided how extensive any validation checks should be, given the number of GEDCOM validators available ([such as this one](http://ged-inline.elasticbeanstalk.com/validate)).
* Work on `tgvis` will begin once an initial version of `tidygedcom` is complete.

## References

* [The GEDCOM Standard - Release 5.5.1](https://edge.fscdn.org/assets/img/documents/ged551-5bac5e57fe88dd37df0e153d9c515335.pdf)
