---
title: "Implementation of tidygedcom"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Why tidygedcom?

One of the main characteristics I wanted for this package was to hide the complexity of the GEDCOM specification. I began with what I wanted the user interface to look like and worked backwards from there. I felt that adopting a **tidyverse** feel to the package would enhance readability and would significantly flatten the learning curve for those who are familiar with **ggplot2**.

## Data structures

I spent a significant amount of time before writing any code considering my options for how the data would be stored under the hood. In [one blog post](https://r-house.netlify.com/2019/12/30/can-genealogical-data-be-tidy/) I considered storing genealogical data in a relational table format as it is easier to deal with, but discounted it very quickly as it is not well suited to nested data (list columns are not ideal for this application).

I toyed with the idea of using an off-the-shelf open source product like [GRAMPS](https://gramps-project.org/blog/) but I found it awkward to use and wanted something where I was in complete control.

I also considered using data structures more suited to this type of data, such as JSON or graphs (using the **igraph** package). However, I discovered it would be quite difficult representing some of the structures in the GEDCOM specification to my satisfaction.

It was some time before a better solution occurred to me, and that was to store the GEDCOM file almost as is, and just to split out the components of each line into their own columns, creating a tidy version of a GEDCOM file (hence the name **tidygedcom**). This would allow easy manipulation using existing **tidyverse** infrastructure, and some additional processing would be needed to deal with the ordered and nested nature of the file. Ironically, I had returned to the idea of a tidy dataframe that I had dismissed early on.

## The tidygedcom object

The **tidygedcom** object is a tibble representation of a GEDCOM file. Before we see how this is structured, let's see an example of a GEDCOM file:

```{r}
readLines(system.file("extdata", "555SAMPLE.GED", package = "tidygedcom"))
```

Lines in a GEDCOM file can have a number of components:

* **Level**: The level in the hierarchical structure. This appears for every line;
* **Cross-reference identifier**: A string (which looks like @XYZ@) that signals the beginning of a new record (apart from header and trailer);
* **Tag**: A short string given immediately after the level or cross-reference identifier that indicates the type of information being provided on the line. These are controlled values. User-defined tags have been allowed in other GEDCOM programs, but they are discouraged here;
* **Cross-reference pointer**: This links to another record in the file (which looks like @XYZ@). In the above example, the Family Group record beginning on line `r which(readLines(system.file("extdata", "555SAMPLE.GED", package = "tidygedcom")) == "0 @F1@ FAM")` references other Individual records who are members of the family;
* **Line value**: The value associated with the tag. For example, on line 6, the CHARacter encoding for the file is given as UTF-8.

There will not be any lines that have all of these components. For example, the first line of records do not contain a line value. Also, a line will never have a cross-reference pointer and a line value defined. For this reason, the **tidygedcom** object treats cross-reference pointers as just another line value.

The **tidygedcom** representation of the above GEDCOM file is given below:

```{r}
library(tidygedcom)

read_gedcom(system.file("extdata", "555SAMPLE.GED", package = "tidygedcom")) %>% 
  print(n = Inf)
```

The cross-reference identifier is given in the record column. Event though these aren't given for the header and trailer records they are assigned as "HD" and "TR" respectively in the **tidygedcom** object.

## File structure

The specifications of GEDCOM records are represented in the `records.R` file. These records are composed of various data structures found in `subrecords.R`. I have also painstakingly written functions to validate each line value separately in `validate_primitives.R` and they are (mostly) all documented in one place in `docs_primitives.R`. These four files form the bulk of the codification of the GEDCOM specification.

The user is kept away from this infrastructure, and instead interacts with various wrapper functions found in the `edit_*.R` files and summary functions found in `interrogate.R`.

## Future work

At the time of writing, this package is getting quite large and difficult to mentally manage. It can be quite tricky implementing some of the functionality to edit existing GEDCOM files (rather than creating from scratch), and so the long term plan is to have a sister package to help with this interactively using **shiny** (**shinygedcom**). However, before work on **shinygedcom** begins, I would like to start a tidygedcom visualisation package, **visgedcom**, which will allow various views of genealogical data.

I am also considering eventually de-scoping **tidygedcom** to only contain the user interface functions and moving the core infrastructure (the four files mentioned above and some other helper functions) to a **tidygedcom.internals** package.

There is also potential to generate another package to handle different types of source documents. With all these package ideas it's turning into a small ecosystem - the **gedcompendium** perhaps.





