---
title: "Cross references and activation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cross references and activation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Cross references

GEDCOM records are given unique identifiers known as xrefs (cross-references) to allow other records to link to them. These are alphanumeric strings surrounded by '@' symbols. The **tidyged** package creates these xrefs automatically:

```{r}
library(tidyged)

simpsons <- gedcom(subm("Me")) %>% 
  add_individual(sex = "M") %>% 
  add_individual_names("Homer /Simpson/") %>% 
  add_individual(sex = "F") %>% 
  add_individual_names("Marge /Simpson/") %>%  
  add_individual(sex = "F") %>% 
  add_individual_names("Lisa /Simpson/") %>% 
  add_individual(sex = "M") %>% 
  add_individual_names("Bart /Simpson/") %>%  
  add_note("This is a note")

dplyr::filter(simpsons, tag %in% c("INDI", "NOTE"))
```

Note the unique xrefs in the record column.

## Activation

In the above example a series of Individual records are created. After each record is created, the name(s) of the individual are defined without actually explicitly referencing the Individual record. This is because they are acting on the *active record*. A record becomes active when it is created or when it is explicitly activated.

We can query the active record using the `get_active_record()` function:

```{r}
get_active_record(simpsons)
```

Since the last record to be created was the Note record, it is the active record. The active record is stored as an attribute of the tibble.

We can use activation to add to existing records. If we want to activate another record, we can activate it using the `activate_*_record()` family of functions together with its xref:

```{r}
simpsons %>% 
  activate_individual_record("@I2@") %>% 
  get_active_record()
```

It can be tedious to have to manually look up these xrefs, and so **tidyged** allows the user to instead enter an expression uniquely identifying the record:

```{r}
simpsons %>% 
  activate_individual_record("Marge") %>% 
  get_active_record()
```

Note that the full name does not need to be given, since the expression is treated as a regular expression. As long as it is detected in the name of the individual it will be found (romanised and phonetic variants of the name are also searched, if they are provided in the record). Note that if forward-slashes are in the name, they should not be used in any expressions.

If no match or more than one match is found, it will result in an error:

```{r, error = TRUE}
simpsons %>% 
  activate_individual_record("Simpson") %>% 
  get_active_record()
```

This functionality can be found when referencing all records, except the Family Group record (where an xref must be provided). The following attributes are used to uniquely identify these records:

* Individuals: Full name (NAME tag), phonetic variation (FONE), romanised variation (ROMN);
* Multimedia: File reference (FILE);
* Note: User text (NOTE);
* Source: Title (TITL);
* Repository: Name (NAME).


