---
title: "Date types"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Date types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

As you would expect with a genealogical data format, dates form a major part of information contained within it. Within `tidyged` there are a number of different date types that can be defined, and there is a helper function for each type to ensure it is formatted correctly.

## Exact dates

Dates that occur on a specific defined date are defined using the `date_exact()` function:

```{r}
library(tidyged)

date_exact(5, 7, 1999)
date_exact(8, 12, 1956)
date_exact(1, 4, 2008)
```

## Date periods

Date periods can take a number of forms. The most intuitive form is when two dates are provided:

```{r}
date_period(start_year = 1999, 
            start_month = 4, 
            start_day = 5, 
            end_year = 2002, 
            end_month = 7, 
            end_day = 12)
```

Not all date components are required. For example:

```{r}
date_period(start_year = 1999, 
            end_year = 2002, 
            end_month = 7)
```

If only the start date is provided, then the return value changes accordingly:

```{r}
date_period(start_year = 1999, 
            start_month = 4, 
            start_day = 5)
```

Alternatively, if you want the date provided to be the end of the period, you can set `to = TRUE`:

```{r}
date_period(start_year = 1999, 
            start_month = 4, 
            start_day = 5,
            to = TRUE)
```

For time periods a very long time ago (where only the year is given), both the `start_year` and/or `end_year` can be interpreted as years before the Common Era:

```{r}
date_period(start_year = 5, 
            end_year = 6,
            start_year_is_bce = TRUE)
```

English dates before 1752 were often given in dual format. This can be achieved using `start_year_is_dual` and `end_year_is_dual`:

```{r}
date_period(start_year = 1746, 
            start_month = 4, 
            start_day = 5, 
            end_year = 1755, 
            end_month = 7, 
            end_day = 12,
            start_year_is_dual = TRUE)
```

## General dates

Sometimes an even more general form of dates are required, and these can be created with the `date_value()` function. This can express both the exact form and date period form described above, as well as:

* Approximate dates
* Date ranges

Approximate dates (i.e. those expressing uncertainty) use the qualifiers 'about', 'calculated', or 'estimated':

```{r}
date_value(start_year = 1999, 
           start_month = 4, 
           start_day = 5, 
           about = TRUE)
date_value(start_year = 1999, 
           start_month = 4, 
           start_day = 5, 
           calc = TRUE)
date_value(start_year = 1999, 
           start_month = 4, 
           start_day = 5, 
           est = TRUE)
```

Date ranges can be controlled using the `before`, `after`, and `between` parameters:

```{r}
date_value(start_year = 1999, 
           start_month = 4, 
           start_day = 5, 
           before = TRUE)
date_value(start_year = 1999, 
           start_month = 4, 
           start_day = 5, 
           after = TRUE)
date_value(start_year = 1999, 
           start_month = 4, 
           start_day = 5, 
           end_year = 2002, 
           end_month = 12, 
           end_day = 25, 
           between = TRUE)
```


