---
title: "Date types"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Date types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

As you would expect with a genealogical data format, dates form a major part of information contained within it. Within `tidyged` there are a number of different date types that can be defined, and there is a helper function for each type to ensure it is formatted correctly.

## Exact dates

Dates that occur on a specific defined date are defined using the `date_exact()` function:

```{r}
library(tidyged)

date_exact(5, 7, 1999)
date_exact(8, 12, 1956)
date_exact(1, 4, 2008)
```

## Calendar dates

Calendar dates are the most common type of date used. They can not only be used to create specific dates like `date_exact()`, they underpin the other date types described below. They can be created using the `date_calendar()` function:

```{r}
date_calendar(year = 1999, month = 4, day = 5)
date_calendar(year = 1999, month = 4)
date_calendar(year = 1999)
date_calendar(month = 4, day = 5)
```

English dates before 1752 were often given in dual format. This can be achieved using the `year_is_dual` parameter:

```{r}
date_calendar(year = 1745, month = 4, day = 5, year_is_dual = TRUE)
```

Years before the Common Era can be defined using the `year_is_bce` parameter:

```{r}
date_calendar(year = 105, year_is_bce = TRUE)
```

Note that only the Gregorian calendar is currently implemented.

## Approximate dates

Approximate dates (i.e. those expressing uncertainty) use the qualifiers 'about', 'calculated', or 'estimated' in conjunction with a `date_calendar()` object:

```{r}
date_calendar(year = 1999, month = 4, day = 5) %>% date_approximated(about = TRUE)
date_calendar(year = 1999, month = 4, day = 5) %>% date_approximated(calc = TRUE)
date_calendar(year = 1999, month = 4, day = 5) %>% date_approximated(est = TRUE)
```


## Date periods

Date periods can take one or two `date_calendar()` objects:

```{r}
date_period(start_date = date_calendar(1956, 7, 26),
            end_date = date_calendar(1956, 9, 15))
date_period(start_date = date_calendar(1956, 7, 26))
date_period(end_date = date_calendar(1956, 9, 15))
```

Providing only one date gives a semi-infinite period.


## Date ranges

Date ranges can be defined using `date_range()`. Unlike `date_period()`, they describe when an event occurred rather than a duration.

```{r}
date_range(start_date = date_calendar(1956, 7, 26),
           end_date = date_calendar(1956, 9, 15))
date_range(start_date = date_calendar(1956, 7, 26))
date_range(end_date = date_calendar(1956, 9, 15))
```


