---
title: "Family group records"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Family group records}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Adding family groups to a tidyged object can be achieved with the `add_famg()` function. The main arguments are about defining the members of the family group. The `husband`, `wife`, and `children` parameters can be xrefs to other Individual records, or partial matches to their names (see `vignette("automation")`). The `number_of_children` parameter describes the total number of known children in this family, regardless of whether they are defined.

The family group records do not need to describe traditional marriage unions - they can describe non-marital relationships and same-sex relationships. Children also do not have to be biological. The parameter names are a legacy from previous versions of the GEDCOM specification, which have not yet been updated.

```{r}
library(tidyged)

simpson_indis <- gedcom(subm("Me")) %>% 
  add_indi(qn = "Homer Simpson", sex = "M") %>% 
  add_indi(qn = "Marge Simpson", sex = "F") %>% 
  add_indi(qn = "Lisa Simpson", sex = "F") %>% 
  add_indi(qn = "Bart Simpson", sex = "M") %>% 
  add_note("These records have not yet been joined in a family group record")

simpson_indis %>% 
  add_famg(husband = "@I1@", wife = "Marge", children = c("Bart", "Lisa"), number_of_children = 3) %>% 
  dplyr::filter(record == "@F1@") %>% 
  knitr::kable()
```

As with other records, various types of user-defined references can be defined (`user_reference_number` and `user_reference_type`), as well as links to multimedia and notes.

## Family events

<p style="text-align: right;"><a href="https://jl5000.github.io/tidyged/articles/repository_records.html">Next article: Repository records ></a></p>
