---
title: "Source records"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Source records}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*WORK IN PROGRESS*

## Introduction

Source records are one of the most important types of record in a GEDCOM file. They provide evidence for the assertions (facts) made in a file, and require careful management.

A Source record can be created with the `add_sour()` function.

```{r}
library(tidyged)

birth_source <- gedcom() |> 
  add_sour(events_recorded = "BIRT, MARR, DEAT",
           date_period_covered = date_period(date_calendar(1900), date_calendar(2000)),
           jurisdiction = "England and Wales",
           originator = "General Register Office",
           title = "Birth Certificate for Joe Bloggs (1956)",
           publication_detail = "something",
           source_text = "This is what is written on the certificate")
```

## Linking back to repositories

```{r}
birth_repo <- birth_source |> 
  add_repo("General Register Office",
           repo_address = address(city = "Southport", state = "Merseyside"))

sour_xref <- find_sour_titl(birth_repo, "Birth Certificate for Joe Bloggs")
repo_xref <- find_repo_name(birth_repo, "General Register Office")

birth_repo |>
  activate_sour(sour_xref) |> 
  add_sour_repo_citation(repo_xref)
```

## Citing source records

## Citation styles

The point to recognise is that users should NOT be exclusively 'splitters' or 'lumpers'.
That choice depends on the Source Citation characteristics.
GRO BMD Index records lend themselves to being Method 2 'lumper' Sources.
BMD Certificates are usually better suited to being Method 1 'splitter' Sources as they often have associated Media images, Text From Source transcripts, and multiple Citations.
So most users adopt a hybrid approach

https://fhug.org.uk/kb/kb-article/citing-sources-method-1-and-method-2/

<p style="text-align: right;"><a href="https://jl5000.github.io/tidyged/articles/multimedia_records.html">Next article: Multimedia records ></a></p>
