---
title: "Summarising tidyged objects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarising tidyged objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `tidyged` package comes with several functions for summarising and querying tidyged objects. The following object is used to illustrate the functionality:

```{r}
library(tidyged)

simpsons <- gedcom(subm("Me"), gedcom_description = "Some of the Simpsons",
                   gedcom_copyright = "No copyright") %>% 
  add_individual("Homer /Simpson/", sex = "M") %>% 
  add_individual_event_birth(event_date = date_calendar(1960, 5, 14),
                             place_name = "Springfield, USA") %>% 
  add_individual("Marge /Simpson/", sex = "F") %>% 
  add_individual_event_birth(event_date = date_calendar(1960, 8, 24),
                             place_name = "Springfield, USA") %>% 
  add_individual("Lisa /Simpson/", sex = "F") %>% 
  add_individual_event_birth(event_date = date_calendar(1984, 12, 3),
                             place_name = "Springfield, USA") %>% 
  add_individual("Bart /Simpson/", sex = "M") %>% 
  add_individual_event_birth(event_date = date_calendar(1982, 3, 16),
                             place_name = "Springfield, USA") %>% 
  add_family_group(husband = "Homer", wife = "Marge", children = c("Bart", "Lisa"), 
                   number_of_children = 3) %>% 
  add_note("This is a note") %>% 
  add_note("This is another note") %>% 
  add_multimedia("123", "PDF") %>% 
  add_multimedia("456", "PDF") %>%
  add_repository("A library") %>% 
  add_source(originator = "J.R. Hartley", title = "A History of the Simpsons") %>%
  add_source(originator = "Prof. J Frink", title = "The screaming and the pain and the hoyvenglaven")
```

## Global summaries

The `str()` and `summary()` functions give a high level overview of tidyged objects:

```{r}
summary(simpsons)
str(simpsons)
```

## Record type summaries

The `num_*()` family of functions count the number of records of a particular type:

```{r}
num_indi(simpsons)
num_fam(simpsons)
num_media(simpsons)
```

These functions are the workhorse of the `str()` function.

The `df_*()` family of functions provide a tabular summary of all records of a particular type:

```{r}
df_individuals(simpsons) %>% knitr::kable()
df_families(simpsons) %>% knitr::kable()
df_multimedia(simpsons) %>% knitr::kable()
```


## Individual record summaries

The `is_*()` family of functions determine whether an xref is a record of a particular type:

```{r}
is_individual(simpsons, "@I1@")
is_family(simpsons, "@F1@")
is_repository(simpsons, "@N1@")
```

The `describe_*()` family of functions give a description of certain records:

```{r}
describe_individual(simpsons, "Bart")
describe_individual(simpsons, "Lisa", name_only = TRUE)
describe_individual(simpsons, "Homer", short_desc = TRUE)
describe_family_group(simpsons, "@F1@")
```

Whilst this family of functions can only operate on one record at a time, the `describe_records()` function can take a whole vector of xrefs, potentially of different types:

```{r}
all_xrefs <- simpsons %>% 
  dplyr::filter(!record %in% c("HD", "TR")) %>% 
  dplyr::pull(record) %>% 
  unique()

describe_records(simpsons, all_xrefs)
```

As with the record-specific functions, setting `short_desc = TRUE` gives a shorter description.

```{r}
describe_records(simpsons, all_xrefs, short_desc = TRUE)
```
