---
title: "Creation of tidyged objects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creation of tidyged objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Importing existing files

The easiest way to create a `tidyged` object is to import an existing GEDCOM file. The `tidyged.io` package (imported with `tidyged`) comes with some sample GEDCOM files, which can be imported using the `read_gedcom()` function:

```{r}
library(tidyged)

tidyged.io::read_gedcom(system.file("extdata", "555SAMPLE.GED", package = "tidyged.io")) %>% 
  print(n = Inf)
```

If you want to export your file as a valid GEDCOM file, you can use the `write_gedcom()` function.

Many other GEDCOM readers will carry out extensive checks on every line of a GEDCOM file ensuring all tags and values are legal, and the grammar is used correctly. The `tidyged` package carries out very few checks, relying on the user to check their files beforehand (as described when loading the package). The few checks that are carried out include:

* Ensuring the file has a valid Byte Order Mark which is consistent with that described in the file;
* Ensuring no lines exceed the character limit;
* Ensuring the header is correctly formed;
* Ensuring there is only one header, trailer, and submitter record defined.

In the future, the `gedcompendium` ecosystem may include a dedicated validation package, but this is currently a low priority.

## Creating new objects

The other way of creating `tidyged` objects is to create them from scratch. A minimal object can be created using an empty `gedcom()` call:

```{r}
gedcom() %>% 
  print(n = Inf)
```

A `tidyged` object actually comes pre-populated with quite a lot of information (more so than a minimal GEDCOM file). Information about the `tidyged` package (as the system/product creating the file) is given in the appropriate places, as well as a default language (English), and submitter details taken from your computer username.

### Submitter

You can use different information for many values using the parameters of the `gedcom()` function. The first argument of the function is a `subm()` object giving details of the submitter:

```{r}
submitter <- subm(name = "The name of the submitter",
                  local_address_lines = "123 Penny Lane Drive",
                  city = "Orlando",
                  state = "Florida",
                  postal_code = "32836",
                  country = "United States of America",
                  phone_number = "123-456-789",
                  email = "email@domain.com",
                  fax = "00000",
                  web_page = "http://www.whousesfaxanymore.com",
                  subm_notes = "The information given above is dummy data")

gedcom(submitter) %>% 
  print(n = Inf)
```

### Other header parameters

The remaining arguments of the `gedcom()` function give further information about the file and the information held within it:

```{r}
gedcom(submitter_details = subm(),
       gedcom_description = "Maternal ancestors of Jamie",
       gedcom_copyright = "No copyright",
       source_data_name = "Some CD I found in the basement",
       source_data_date = date_exact(12, 5, 2009),
       source_data_copyright = "No data copyright",
       receiving_system = "tidyged",
       language = "English") %>% 
  print(n = Inf)
```

Note that:

* The `gedcom_description` parameter should describe the type of genealogical data contained in the file;
* The date associated with the source data must be given as a `date_exact()` object to ensure it is formatted correctly (see `vignette("dates")` for more information on dates);
* The `receiving_system` should be left to its default value, unless you are specifically intending on passing it to a different system.

